On the adhesivity of EGGS

### 2025-02-13 calco/main.tex
minori
Update main.tex
conclusioni

\title{On the adhesivity properties of equivalence graphs\thanks{???????}}

@@

\tikzset{->-/.style={decoration={
			markings,
			mark=at position #1 with {\arrow{>}}},postaction={decorate}}}
	\maketitle \todo{Scegliere un titolo vero e mettere i ringraziamenti}
	\begin{abstract}
\todo{a very nice abstract}
	\end{abstract}

@@

\subsection{Hierarchical hypergraphs}

\todo{da fare? lasciare per una versione estesa?}

@@

\section{Conclusions and further works}
\todo{Some very nice conclusions}
\bibliography{biblio}

>>>>

\title{On the adhesivity of EGGS}

@@

\tikzset{->-/.style={decoration={
			markings,
			mark=at position #1 with {\arrow{>}}},postaction={decorate}}}

\maketitle

\begin{abstract}
\todo{a very nice abstract}
\end{abstract}

@@

@@

\section{Conclusions and further works}
The aim of our paper was to extend the theory of adhesive categories in order to include EGGS, 
an up-and-coming formalism for program optimisation and synthesis via a compact representation and 
efficient implementation of equality saturation.
%
To do so, we revisited and generalised the notions of hyper-graph and term graph with equivalence, and
we extended it in order to capture EGGS as term graphs satisfying a suitable closure property.

Our result opens two threads of research. The first is to use the quasi-adhesivity of EGGS in order to
model their rewriting via the double-pushout (DPO) approach. This is likely to be straightforward, since 
the rules adopted in the literature of EGGS seems to be span of regular monos, and such rules
perfectly fit the mold of rewriting on quasi-adhesive categories. 
For example, the equivalence $x \div x = 1$, taken from the introductory example in~\cite{xxx}, 
can be modelled as the rule
\[to draw DPO rule\]

It still needs to be investigated what concurrency and termination, the key properties for DPO rewriting
on adhesive categories, 
means in the context of EGGS. More interestingly, another venue for development is using the adhesive 
machinery to extend the EGGS formalism. In fact, most of the results
presented here for hyper-graphs can be generalised to hierarchical hyper-graphs, that is, 
hypergraphs with a hierarchy (a partial order) among 
edges that is useful for adding structural information, such as encapsulation and sandboxing~\cite{xxx}.

Finally, we will delve into and draw a comparison with an alternative categorical presentation for EGGS 
advanced in~\cite{ghica}. The proposal is quite different from our own. Simplifying, the key is to use categories 
of tree which are equipped with a lattice on hom-sets. We can recover trees in our approach, even 
if we opted for term graphs in order to obtain a simpler categorical presentation. Instead of going through 
an enrichment on hom-sets, we chose to obtain closure via suitable requirements on morphisms, even if it 
seems that the former generalises the latter, at the expenses of a more complex machinery.

\bibliography{biblio}
